Generalized Parallel Computational Schemes\\ for Time-Consuming Global Optimization

### 2018-01-18 2017/KAZAN_LJM/paper_lobach.tex
some corrections

\begin{document}
\titlerunning{Generalized Parallel Schemes for Global Optimization}
\authorrunning{Strongin, Gergel, Barkalov, Sysoyev}

@@

\author{\firstname{R.~G.}~\surname{Strongin}}
\email[E-mail: ]{president@unn.ru}
\affiliation{Lobachevsky State University of Nizhni Novgorod}

\author{\firstname{V.~P.}~\surname{Gergel}}
\email[E-mail: ]{gergel@unn.ru}
\affiliation{Lobachevsky State University of Nizhni Novgorod}

\author{\firstname{K.~A.}~\surname{Barkalov}}
\email[E-mail: ]{barkalov@vmk.unn.ru}
\affiliation{Lobachevsky State University of Nizhni Novgorod}

\author{\firstname{A.~V.}~\surname{Sysoyev}}
\email[E-mail: ]{sysoyev@vmk.unn.ru}
\affiliation{Lobachevsky State University of Nizhni Novgorod}

@@

The state-of-the-art in the field of global optimization is outlined in a number of key works – see, for instance, \cite{Floudas, Locatelli, Strongin1, Pardalos, Sergeyev1, Paulavicius}. The development of the theory and methods of use of the high-performance computing for solving the computationally intensive problems of global optimization is one of the trends developing rapidly \cite{Strongin1, Ciegis, Luque, Strongin2}.

@@

In the framework of the information-statistical theory, the dimensionality reduction is provided by using the Peano space-filling curves $y(x)$, which map the interval $[0, 1]$ onto a $N$-dimensional domain $D$ \cite{Strongin1, Sergeyev2, Strongin3}. As a result of such reduction, the initial multidimensional global optimization problem (1) is reduced to a one-dimensional one:
\begin{equation}
\varphi(y(x^\ast)) = \min\{\varphi(y(x)): x \in [0,1]\}.
\end{equation}
for which the function $\varphi(y(x))$ satisfies a uniform H\"older condition \cite{Strongin1}, i.e.
\begin{equation}
|\varphi(y(x')-\varphi(y(x''))| \leq H|x' - x''|^{1/N}, x', x'' \in [0,1]\},
\end{equation}
where the H\"older constant $H$ is determined as $H = 2L\sqrt{N + 3}$, $L$ is the Lipschitz constant from (2) and $N$ is the optimization problem dimension (1). A one-dimensional function $\varphi(y(x))$ is produced by reduction of the multi-dimensional function $\varphi(y)$. In order to calculate the values of $\varphi(y(x))$, the multidimensional image $y \in D$ should be computed first in accordance with the mapping $y(x)$, and then the value of the multidimensional function $\varphi(y)$ at the multidimensional point $y \in D$ should be evaluated. The resulting value $z = \varphi(y)$ is used further as the value of the one-dimensional function $\varphi(y(x))$.

@@

It should be noted that the proposed approach can also be applied for solving many other computationally intensive problems -- such as root finding in systems of nonlinear equations, integration of multidimensional integrals, approximation of the data dependencies etc.

@@

The stopping condition to terminate the optimization iterations is defined by the inequality
\begin{equation}
(x_t - x_{t-1}) \leq \varepsilon,
\end{equation}
for the interval $t, 1 \leq i \leq k + 1$ with the maximum characteristic from (6) and $\varepsilon > 0$ is the required problem solving accuracy. If the stopping condition is not satisfied, the iteration number $k$ is incremented by unity, and a new global search iteration is executed.

In order to explain the computational scheme given above, the following can be noted. The computed characteristics $R(i), 1 \leq i \leq k + 1$ can be understood as the importance indicators for the intervals with respect to the global minimum point location. This provides the explanation for the interval selection scheme for executing the next trial: the point of every next global search iteration is selected in the interval with the maximum characteristic (i.e. in the interval, which is the most likely one to contain the global minimum point).

According to this general scheme, specific global search algorithms are formulated by defining the formulas to compute the characteristics $R(i), 1 \leq i \leq k+1$, and the points of the next trials in the intervals with maximum characteristics. Thus, for a multidimensional general algorithm of global search (MGAGS) \cite{Strongin1}, the interval characteristic is defined as
\begin{equation}
R(i)=m\rho_i+\frac{(z_i-z_{i-1})^2}{m\rho_i}-2 (z_i + z_{i-1}), 1 \leq i \leq k + 1,
\end{equation}
where $m$ is the H\"older constant estimate from (4), $z_i = \varphi(y(x_i)), 1 \leq i \leq k + 1$ are the computed values of the minimized function $\varphi(y(x))$ at the points of the executed iterations, and $\rho_i = \left(x_i - x_{i-1} \right)^{1/N}, 1 \leq i \leq k + 1$.

The next trial point in the case of MGAGS is computed as follows:
\begin{equation}
x^{k+1} = \frac{x_t+x_{t-1}}{2} - \mathrm{sign}(z_t-z_{t-1})\frac{1}{2r}\left[\frac{\left|z_t-z_{t-1}\right|}{m}\right]^N,
\end{equation}
where $r, r > 1$ is a \textit{parameter} of the algorithm.

@@

As it has been noted above, the general approach to provide the parallel computations consists in simultaneous calculating the minimized function values at several different points of the searching domain $D$. For the shared memory computing systems, selecting the set of points for executing the parallel trials can be based on the characteristics $R(i), 1 \leq i \leq k + 1$ of the search intervals $(x_{i-1} ,x_i), 1 \leq i \leq k + 1$ introduced in the framework of the general computational scheme (see Subsection 3.1). As the characteristics are understood as the importance indicators for the intervals with respect to the global minimum point location, additional intervals for several trials to be run in parallel should be selected subject to the following values of the characteristics.

As a result, a general scheme of the parallel algorithms of global optimization for systems with shared memory can be formulated on the basis of the scheme given in Subsection 3.1 where the following set of rules should be modified (see also \cite{Strongin1, Strongin2, Strongin4, Grishagin2}):

\textit{Rule} $3'$. Arrange the interval characteristics in the descending order
\begin{equation}
R(t_1) \geq R(t_2) \geq \dots \geq R(t_k) \geq R(t_{k+1})
\end{equation}
and select $p$ intervals with the numbers $t_j, 1 \leq j \leq p$ having the maximum values of characteristics ($p$ is the number of processors (cores) used for the parallel computing).

\textit{Rule} $4'$. Execute new trials (compute the values of the minimized function $\varphi(y(x))$) at the points $x^{k+j}, 1 \leq j \leq p$ located within the intervals with the maximum characteristics from (10).

@@

In general, the described general scheme of parallel algorithms of global optimization for systems with shared memory identifies the procedure for competitive parallel computations (see Fig.~4a) where the search intervals $(x_{i-1} ,x_i), 1 \leq i \leq k + 1$ compete with each other for the access to the processors to execute next global search iterations. Taking into account the possible number of processors/cores in computational nodes with homogeneous shared memory, the described approach enables the parallelism of about $10^2$ computing units.

@@

Generating such a family of problems can take place during the solving of the multicriterial optimization problems -- see, for example \cite{Gergel3, Gergel4}. Another approach used widely was proposed in \cite{Strongin1, Strongin2, Strongin5, Gergel1}. This approach uses a set of mappings
\begin{equation}
Y_s(x) = \{ y^1(x), \dots , y^s(x) \},
\end{equation}
allowing constructing the necessary family of optimization problems 
\begin{equation}
\phi\bigl(y^l(x^\ast_l)\bigr) = \min \bigl\{\phi\bigl(y^l(x)\bigl): x \in [0, 1] \bigl\}, 1 \leq l \leq s,
\end{equation}
for which the global minimum points coincide with the solution of the initial problem (1), i.e. $ y^\ast = y^l(x_l^\ast), 1 \leq l \leq s$. It is important to note that a family of the one-dimensional problems $\phi(y^l(x)), 1 \leq l \leq s$ resulting from the dimensionality reduction is an information-linked one: the function values computed for some function $\phi(yl(x))$ can be used by all other problems of the family (14).

@@

\begin{itemize}
\item Before starting a new trial for any problem $\phi(y^l(x)), 1 \leq l \leq s$ at any point $x' \in [0,1]$, it is necessary to calculate the image $y' \in D$ of the point $x' \in [0,1]$ according to the mapping of $y^l(x)$ and compute the preimages of $x_i', 1 \leq i \leq s$ for each problem of the family $\overrightarrow{\Phi}(y)$. The resulting preimages $x_i', 1 \leq i \leq s$ should be sent to all involved computational nodes to eliminate repeated selection of new trial points in the intervals containing the preimages.

\item Upon the completing of any trial for any problem $\phi(y^l(x)), 1 \leq l \leq s$ at any point $x' \in [0,1]$, calculate all preimages $x_i', 1 \leq i \leq s$ again and distribute these preimages and the trial result $z' = \phi(y^l(x'))$ among all involved computational nodes to include the obtained data in the search information (5).

@@

The multistage scheme (15) can be generalized for the integration with the dimensionality reduction scheme based on the Peano curves \cite{Sysoyev}. According to the generalized multistage block scheme, the vector of variables $y \in D$ of the global optimization problem (1) is considered to be a set of block variables
\[
y = (y_1, y_2, ..., y_N) = (u_1, u_2, ..., u_M)
\]
where the $i$-th block variable $u_i$ consists of the consecutive elements of the vector $y$, i.e.
\[
u_1 = (y_1, y_2, ..., y_{N_1}), \dots, 
u_M = (y_{N-N_M+1}, y_{N-N_M+2}, ..., y_N).
\]
where $N_1 + N_2 + \dots + N_M = N$.

@@

The computational experiments were performed using the Lobachevsky supercomputer at the State University of Nizhny Novgorod (operating system -- CentOS 6.4, supercomputer management system - SLURM). Each computational node had two 8-core Intel Sandy Bridge E5-2660 2.2 GHz CPUs, 64 GB RAM, and 3 NVIDIA Kepler K20Х graphic processors ($2\,688$ cores, 6 GB GDDR5 RAM). Intel C++ 14.0.2 was used to generate the executable program code. SDK CUDA Toolkit 6.0 was used to generate the applications for the graphic processors.

@@

To obtain more reliable estimates of the parallelization efficiency, 100 test problems of global optimization were solved. The global minima $y^\ast$ of the test problems were considered to be found if a trial point $y^k$ is generated in the $\delta$-vicinity of the global minimum, i.e. $\| y^k - y^ \ast \| \leq \delta$. The vicinity size was set equal to $\delta = 0.01 \|b - a \|$ , where $a$ and $b$ were the boundaries of the search domain $D$. The maximum allowed number of parallel iterations was $Kmax = 10\,000$.

The computational experiments involved the use of $p$ nodes of the Lobachevsky supercomputer, each node used three NVIDIA Kepler K20Х graphic accelerators. Thus, for $p = 32$, the number of GPU accelerators were 96, each having $2\,688$ CUDA cores (i.e. total $258\,048$ CUDA cores). Table 1 shows the Globalizer execution time, while Table 2 gives obtained parallel speedup with respect to the execution time for single computational node.

@@

\begin{acknowledgments}
This research was supported by the Russian Science Foundation, project No 16-11-10150 "Novel efficient methods and software tools for time-consuming decision making problems using supercomputers of superior performance".
\end{acknowledgments}

@@

\bibitem{Strongin4}
R.~G.~Strongin and Y.~D.~Sergeyev, \textit{Global multidimensional optimization on parallel computer}, Parallel Computing, \textbf{18} (11), 1259--1273, (1992).

\bibitem{Sergeyev4}
Y.~D.~Sergeyev and V.~A.~Grishagin, \textit{Parallel asynchronous global search and the nested optimization scheme}, J. Comput. Anal. Appl., \textbf{3} (2), 123--145, (2001).

\bibitem{Gergel1}
V.~P.~Gergel and S.~V.~Sidorov, \textit{A Two-Level Parallel Global Search Algorithm for Solution of Computationally Intensive Multiextremal Optimization Problems}, V. Malyshkin (Ed.): PaCT 2015, LNCS 9251, 505--515. Springer-Verlag Berlin Heidelberg ,(2015).

\bibitem{Gergel2}
V.~Gergel, \textit{An Unified Approach to Use of Coprocessors of Various Types for Solving Global Optimization Problems}, 2nd International Conference on Mathematics and Computers in Sciences and in Industry, 13--18, (2015).

\bibitem{Barkalov}
K.~Barkalov, V.~Gergel and I.~Lebedev, \textit{Solving global optimization problems on GPU cluster}, In: Simos, T.E. (ed.) ICNAAM 2015. AIP Conference Proceedings, 1738, art. no. 400006, (2016).

\bibitem{Gergel3}
V.~Gergel and E.~Kozinov, \textit{Efficient methods of multicriterial optimization based on the intensive use of search information}, Springer Proceedings in Mathematics and Statistics, 197, 27--45, (2017).

\bibitem{Gergel4}
V.~Gergel and E.~Kozinov, \textit{Parallel computing for time-consuming multicriterial optimization problems}, Lecture Notes in Computer Science, 10421, 446--458, (2017).

\bibitem{Gergel5}
V.~Gergel, V.~Grishagin and A.~Gergel, \textit{Adaptive nested optimization scheme for multidimensional global search}, Journal of Global Optimization, \textbf{66} (1), 35--51, (2016).

\bibitem{Lera}
D.~Lera, Y.~D.~Sergeyev, \textit{Lipschitz and Holder global optimization using space-filling curves}, Appl. Numer. Math, \textbf{60} (1-2), 115--129, (2010).

\bibitem{Grishagin1}
V~A.~Grishagin, \textit{On convergence conditions for a class of global search algorithms}, Proceedings of the 3-rd All-Union seminar "Numerical methods of nonlinear programming", Kharkov, 82--84, (1979) [In Russian].

\bibitem{Grishagin2}
V.~A.~Grishagin, Y.~D.~Sergeyev and R.~G.~Strongin, \textit{Parallel characteristic algorithms for solving problems of global optimization}, Journal of Global Optimization, \textbf{10} (2), 185--206, (1997).

\bibitem{Strongin5}
R.~G.~Strongin, \textit{Algorithms for Multi-extremal Mathematical Programming Problems Employing the Set of Joint Space-filling Curves}, Journal of Global Optimization, \textbf{2} (4), 357--378, (1992).

\bibitem{Sysoyev}
A.~Sysoyev, K.~Barkalov, V.~Sovrasov, I.~Lebedev and V.~Gergel, \textit{Globalizer –- A parallel software system for solving global optimization problems}, Lecture Notes in Computer Science 10421, 492--499, (2017).

\bibitem{Sergeyev7}
Y.~D.~Sergeyev and V.~A.~Grishagin, \textit{Parallel Asynchronous Global Search and the Nested Optimization Scheme}, J. Comput. Anal. Appl., \textbf{3} (2), 123--145, (2001).

\bibitem{Gaviano}
M.~Gaviano, D.~Lera, D.~E.~Kvasov and Ya.~D.~Sergeyev, \textit{Software for generation of classes of test functions with known local and global minima for global optimization}, ACM Trans. Math. Software, \textbf{29}, 469--480, (2003).

\bibitem{Modorskii}
V.~Y.~Modorskii, D.~F.~Gaynutdinova, V.~P.~Gergel and K.~A.~Barkalov, \textit{Optimization in design of scientific products for purposes of cavitation problems}, AIP Conference Proceedings, 1738, art. no. 400013, (2016).

\bibitem{Gergel6}
V.~P.~Gergel, M.~I.~Kuzmin, N.~A.~Solovyov and V.~A.~Grishagin, \textit{Recognition of surface defects of cold-rolling sheets based on method of localities}, International Review of Automatic Control, \textbf{8} (1), 51--55, (2015).

>>>>

\begin{document}
\titlerunning{Generalized Parallel Schemes for Global Optimization}
\authorrunning{Strongin et al.}

@@

\author{\firstname{R.~G.}~\surname{Strongin}}
\email[E-mail: ]{president@unn.ru}
\affiliation{Lobachevsky State University of Nizhni Novgorod, pr. Gagarina 23, Nizhni Novgorod, Russia}

\author{\firstname{V.~P.}~\surname{Gergel}}
\email[E-mail: ]{gergel@unn.ru}
\affiliation{Lobachevsky State University of Nizhni Novgorod, pr. Gagarina 23, Nizhni Novgorod, Russia}

\author{\firstname{K.~A.}~\surname{Barkalov}}
\email[E-mail: ]{barkalov@vmk.unn.ru}
\affiliation{Lobachevsky State University of Nizhni Novgorod, pr. Gagarina 23, Nizhni Novgorod, Russia}

\author{\firstname{A.~V.}~\surname{Sysoyev}}
\email[E-mail: ]{sysoyev@vmk.unn.ru}
\affiliation{Lobachevsky State University of Nizhni Novgorod, pr. Gagarina 23, Nizhni Novgorod, Russia}

@@

The state-of-the-art in the field of global optimization is outlined in a number of key works -- see, for instance, \cite{Floudas, Locatelli, Strongin1, Pardalos, Sergeyev1, Paulavicius}. The development of the theory and methods of use of the high-performance computing for solving the computationally intensive problems of global optimization is one of the trends developing rapidly \cite{Strongin1, Ciegis, Luque, Strongin2}.

@@

In the framework of the information-statistical theory, the dimensionality reduction is provided by using the Peano space-filling curves $y(x)$, which map the interval $[0, 1]$ onto a $N$-dimensional domain $D$ \cite{Strongin1, Sergeyev2, Strongin3}. As a result of such reduction, the initial multidimensional global optimization problem (1) is reduced to a one-dimensional one 
\begin{equation}
\varphi(y(x^\ast)) = \min\{\varphi(y(x)): x \in [0,1]\},
\end{equation}
for which the function $\varphi(y(x))$ satisfies a uniform H\"older condition \cite{Strongin1}, i.e.
\begin{equation}
|\varphi(y(x')-\varphi(y(x''))| \leq H|x' - x''|^{1/N}, x', x'' \in [0,1]\},
\end{equation}
where the H\"older constant $H$ is determined as $H = 2L\sqrt{N + 3}$, $L$ is the Lipschitz constant from (2) and $N$ is the optimization problem dimension (1). A one-dimensional function $\varphi(y(x))$ is produced by reduction of the multi-dimensional function $\varphi(y)$. In order to calculate the values of $\varphi(y(x))$, the multidimensional image $y \in D$ should be computed first in accordance with the mapping $y(x)$, and then the value of the multidimensional function $\varphi(y)$ at the multidimensional point $y \in D$ should be evaluated. The resulting value $z = \varphi(y)$ is used further as the value of the one-dimensional function $\varphi(y(x))$.

@@

It should be noted that the proposed approach can also be applied for solving many other computationally intensive problems such as root finding in systems of nonlinear equations, integration of multidimensional integrals, approximation of the data dependencies etc.

@@

The stopping condition to terminate the optimization iterations is defined by the inequality
\begin{equation}
(x_t - x_{t-1}) \leq \varepsilon
\end{equation}
for the interval $t$ with the maximum characteristic from (6); $\varepsilon > 0$ is the required problem solving accuracy. If the stopping condition is not satisfied, the iteration number $k$ is incremented by unity, and a new global search iteration is executed.

In order to explain the computational scheme given above, the following can be noted. The computed characteristics $R(i), 1 \leq i \leq k + 1$, can be understood as the importance indicators for the intervals with respect to the global minimum point location. This provides the explanation for the interval selection scheme for executing the next trial: the point of every next global search iteration is selected in the interval with the maximum characteristic (i.e. in the interval, which is the most likely one to contain the global minimum point).

According to this general scheme, specific global search algorithms are formulated by defining the formulas to compute the characteristics $R(i), 1 \leq i \leq k+1$, and the points of the next trials in the intervals with maximum characteristics. Thus, for a multidimensional general algorithm of global search (MGAGS) \cite{Strongin1}, the interval characteristic is defined as
\begin{equation}
R(i)=m\rho_i+\frac{(z_i-z_{i-1})^2}{m\rho_i}-2 (z_i + z_{i-1}), 1 \leq i \leq k + 1,
\end{equation}
where $m$ is the H\"older constant estimate from (4), $z_i = \varphi(y(x_i)), 1 \leq i \leq k + 1$, are the computed values of the minimized function $\varphi(y(x))$ at the points of the executed iterations, and $\rho_i = \left(x_i - x_{i-1} \right)^{1/N}, 1 \leq i \leq k + 1$.

The next trial point in the case of MGAGS is computed as follows:
\begin{equation}
x^{k+1} = \frac{x_t+x_{t-1}}{2} - \mathrm{sign}(z_t-z_{t-1})\frac{1}{2r}\left[\frac{\left|z_t-z_{t-1}\right|}{m}\right]^N,
\end{equation}
where $r > 1$ is a \textit{parameter} of the algorithm.

@@

As it has been noted above, the general approach to provide the parallel computations consists in simultaneous calculating the minimized function values at several different points of the searching domain $D$. For the shared memory computing systems, selecting the set of points for executing the parallel trials can be based on the characteristics $R(i), 1 \leq i \leq k + 1$, of the search intervals $(x_{i-1} ,x_i), 1 \leq i \leq k + 1$, introduced in the framework of the general computational scheme (see Subsection 3.1). As the characteristics are understood as the importance indicators for the intervals with respect to the global minimum point location, additional intervals for several trials to be run in parallel should be selected subject to the following values of the characteristics.

As a result, a general scheme of the parallel algorithms of global optimization for systems with shared memory can be formulated on the basis of the scheme given in Subsection 3.1 where the following set of rules should be modified (see also \cite{Strongin1, Strongin2, Strongin4, Grishagin2}).

\textit{Rule} $3'$. Arrange the interval characteristics in the descending order
\begin{equation}
R(t_1) \geq R(t_2) \geq \dots \geq R(t_k) \geq R(t_{k+1})
\end{equation}
and select $p$ intervals with the numbers $t_j, 1 \leq j \leq p$, having the maximum values of characteristics ($p$ is the number of processors (cores) used for the parallel computing).

\textit{Rule} $4'$. Execute new trials (compute the values of the minimized function $\varphi(y(x))$) at the points $x^{k+j}, 1 \leq j \leq p$, located within the intervals with the maximum characteristics from (10).

@@

In general, the described general scheme of parallel algorithms of global optimization for systems with shared memory identifies the procedure for competitive parallel computations (see Fig.~4a) where the search intervals $(x_{i-1} ,x_i), 1 \leq i \leq k + 1$, compete with each other for the access to the processors to execute next global search iterations. Taking into account the possible number of processors/cores in computational nodes with homogeneous shared memory, the described approach enables the parallelism of about $10^2$ computing units.

@@

Generating such a family of problems can take place during the solving of the multicriterial optimization problems -- see, for example \cite{Gergel3, Gergel4}. Another approach used widely was proposed in \cite{Strongin1, Strongin2, Strongin5, Gergel1}. This approach uses a set of mappings
\begin{equation}
Y_s(x) = \{ y^1(x), \dots , y^s(x) \}
\end{equation}
allowing constructing the necessary family of optimization problems 
\begin{equation}
\phi\bigl(y^l(x^\ast_l)\bigr) = \min \bigl\{\phi\bigl(y^l(x)\bigl): x \in [0, 1] \bigl\}, 1 \leq l \leq s,
\end{equation}
for which the global minimum points coincide with the solution of the initial problem (1), i.e. $ y^\ast = y^l(x_l^\ast), 1 \leq l \leq s$. It is important to note that a family of the one-dimensional problems $\phi(y^l(x)), 1 \leq l \leq s$, resulting from the dimensionality reduction is an information-linked one: the function values computed for some function $\phi(yl(x))$ can be used by all other problems of the family (14).

@@

\begin{itemize}
\item Before starting a new trial for any problem $\phi(y^l(x)), 1 \leq l \leq s$, at any point $x' \in [0,1]$, it is necessary to calculate the image $y' \in D$ of the point $x' \in [0,1]$ according to the mapping of $y^l(x)$ and compute the preimages of $x_i', 1 \leq i \leq s$, for each problem of the family $\overrightarrow{\Phi}(y)$. The resulting preimages $x_i', 1 \leq i \leq s$, should be sent to all involved computational nodes to eliminate repeated selection of new trial points in the intervals containing the preimages.

\item Upon the completing of any trial for any problem $\phi(y^l(x)), 1 \leq l \leq s$, at any point $x' \in [0,1]$, calculate all preimages $x_i', 1 \leq i \leq s$, again and distribute these preimages and the trial result $z' = \phi(y^l(x'))$ among all involved computational nodes to include the obtained data in the search information (5).

@@

The multistage scheme (15) can be generalized for the integration with the dimensionality reduction scheme based on the Peano curves \cite{Sysoyev}. According to the generalized multistage block scheme, the vector of variables $y \in D$ of the global optimization problem (1) is considered to be a set of block variables
\[
y = (y_1, y_2, ..., y_N) = (u_1, u_2, ..., u_M),
\]
where the $i$-th block variable $u_i$ consists of the consecutive elements of the vector $y$, i.e.
\[
u_1 = (y_1, y_2, ..., y_{N_1}), \dots, 
u_M = (y_{N-N_M+1}, y_{N-N_M+2}, ..., y_N),
\]
where $N_1 + N_2 + \dots + N_M = N$.

@@

The computational experiments were performed using the Lobachevsky supercomputer at the State University of Nizhni Novgorod (CentOS 6.4 operating system, SLURM supercomputer management system). Each computational node had two 8-core Intel Sandy Bridge E5-2660 2.2 GHz CPUs, 64 GB RAM, and 3 NVIDIA Kepler K20X graphic processors ($2\,688$ cores, 6 GB GDDR5 RAM). Intel C++ 14.0.2 was used to generate the executable program code. SDK CUDA Toolkit 6.0 was used to generate the applications for the graphic processors.

@@

To obtain more reliable estimates of the parallelization efficiency, 100 test problems of global optimization were solved. The global minima $y^\ast$ of the test problems were considered to be found if a trial point $y^k$ is generated in the $\delta$-vicinity of the global minimum, i.e. $\| y^k - y^ \ast \| \leq \delta$. The vicinity size was set equal to $\delta = 0.01 \|b - a \|$ , where $a$ and $b$ were the boundaries of the search domain $D$. The maximum allowed number of parallel iterations was $K_{max} = 10\,000$.

The computational experiments involved the use of $p$ nodes of the Lobachevsky supercomputer, each node used three NVIDIA Kepler K20X graphic accelerators. Thus, for $p = 32$, the number of GPU accelerators were 96, each having $2\,688$ CUDA cores (i.e. total $258\,048$ CUDA cores). Table 1 shows the Globalizer execution time, while Table 2 gives obtained parallel speedup with respect to the execution time for single computational node.

@@

\begin{acknowledgments}
This research was supported by the Russian Science Foundation, project No 16-11-10150 ``Novel efficient methods and software tools for time-consuming decision making problems using supercomputers of superior performance''.
\end{acknowledgments}

@@

\bibitem{Strongin4}
R.~G.~Strongin and Y.~D.~Sergeyev, \textit{Global multidimensional optimization on parallel computer}, Parallel Computing \textbf{18} (11), 1259--1273 (1992).

\bibitem{Sergeyev4}
Y.~D.~Sergeyev and V.~A.~Grishagin, \textit{Parallel asynchronous global search and the nested optimization scheme}, J. Comput. Anal. Appl., \textbf{3} (2), 123--145 (2001).

\bibitem{Gergel1}
V.~P.~Gergel and S.~V.~Sidorov, \textit{A Two-Level Parallel Global Search Algorithm for Solution of Computationally Intensive Multiextremal Optimization Problems}, Lecture Notes in Computer Science \textbf{9251}, 505--515 (2015).

\bibitem{Gergel2}
V.~Gergel, \textit{An Unified Approach to Use of Coprocessors of Various Types for Solving Global Optimization Problems}, in Proceedings of the Second International Conference on Mathematics and Computers in Sciences and in Industry (MCSI), Sliema, 13--18 (2015).

\bibitem{Barkalov}
K.~Barkalov, V.~Gergel and I.~Lebedev, \textit{Solving global optimization problems on GPU cluster}, AIP Conference Proceedings \textbf{1738}, 400006 (2016).

\bibitem{Gergel3}
V.~Gergel and E.~Kozinov, \textit{Efficient methods of multicriterial optimization based on the intensive use of search information}, Springer Proceedings in Mathematics and Statistics \textbf{197}, 27--45 (2017).

\bibitem{Gergel4}
V.~Gergel and E.~Kozinov, \textit{Parallel computing for time-consuming multicriterial optimization problems}, Lecture Notes in Computer Science \textbf{10421}, 446--458 (2017).

\bibitem{Gergel5}
V.~Gergel, V.~Grishagin and A.~Gergel, \textit{Adaptive nested optimization scheme for multidimensional global search}, J. Glob. Optim.  \textbf{66} (1), 35--51 (2016).

\bibitem{Lera}
D.~Lera, Y.~D.~Sergeyev, \textit{Lipschitz and Holder global optimization using space-filling curves},  Appl. Numer. Math. \textbf{60} (1-2), 115--129 (2010).

\bibitem{Grishagin1}
V~A.~Grishagin. \textit{On convergence conditions for a class of global search algorithms}, in Proceedings of the 3-rd All-Union seminar ``Numerical methods of nonlinear programming'', Kharkov, 82--84 (1979) [In Russian].

\bibitem{Grishagin2}
V.~A.~Grishagin, Y.~D.~Sergeyev and R.~G.~Strongin, \textit{Parallel characteristic algorithms for solving problems of global optimization}, J. Glob. Optim. \textbf{10} (2), 185--206 (1997).

\bibitem{Strongin5}
R.~G.~Strongin, \textit{Algorithms for Multi-extremal Mathematical Programming Problems Employing the Set of Joint Space-filling Curves}, J. Glob. Optim. \textbf{2} (4), 357--378 (1992).

\bibitem{Sysoyev}
A.~Sysoyev, K.~Barkalov, V.~Sovrasov, I.~Lebedev and V.~Gergel, \textit{Globalizer -- A parallel software system for solving global optimization problems}, Lecture Notes in Computer Science \textbf{10421}, 492--499 (2017).

\bibitem{Sergeyev7}
Y.~D.~Sergeyev and V.~A.~Grishagin, \textit{Parallel Asynchronous Global Search and the Nested Optimization Scheme}, J. Comput. Anal. Appl. \textbf{3} (2), 123--145 (2001).

\bibitem{Gaviano}
M.~Gaviano, D.~Lera, D.~E.~Kvasov and Ya.~D.~Sergeyev, \textit{Software for generation of classes of test functions with known local and global minima for global optimization}, ACM Trans. Math. Softw. \textbf{29}, 469--480 (2003).

\bibitem{Modorskii}
V.~Y.~Modorskii, D.~F.~Gaynutdinova, V.~P.~Gergel and K.~A.~Barkalov, \textit{Optimization in design of scientific products for purposes of cavitation problems}, AIP Conference Proceedings \textbf{1738}, 400013 (2016).

\bibitem{Gergel6}
V.~P.~Gergel, M.~I.~Kuzmin, N.~A.~Solovyov and V.~A.~Grishagin, \textit{Recognition of surface defects of cold-rolling sheets based on method of localities}, International Review of Automatic Control \textbf{8} (1), 51--55 (2015).
